---
title: "Isotopes_bear_hair"
author: "Nere"
date: "10/24/2019"
output: html_document
---

OPEN AND REVIEW 
```{r, set.seed(123)}
knitr::opts_chunk$set(cache = T)

hair <- read.csv(file="/Users/nereyda/Documents/isotopos/hair_N15.csv", header=TRUE, sep=",")
summary(hair)
```


```{r}
library(lattice) 
library(ggplot2)
library(tidyr)
library(dplyr)
library(R2jags)
```

Adding a ID to species for the graph
```{r}
hair$speciesID<-hair$species2
levels(hair$speciesID) <- c(levels(hair$speciesID), "AM","CE","CL","CT","EB","EC","LR","MU","MF","MM","Ma","OV","PB","PH","PL","PV","RT","RU","TOP","TOV","UA","UR","HM","UM","UT","VV")
hair$speciesID[hair$speciesID =='A.melanoleuca']<-'AM'
hair$speciesID[hair$speciesID =='C.elaphus']<-'CE'
hair$speciesID[hair$speciesID =='C.lupus']<-'CL'
hair$speciesID[hair$speciesID =='C.temminckii']<-'CT'
hair$speciesID[hair$speciesID =='E.Barbara']<-'EB'
hair$speciesID[hair$speciesID =='E.cephalopus']<-'EC'
hair$speciesID[hair$speciesID =='L.rufus']<-'LR'
hair$speciesID[hair$speciesID =='M. ursinus']<-'MU'
hair$speciesID[hair$speciesID =='M.foina']<-'MF'
hair$speciesID[hair$speciesID =='M.moschata']<-'MM'
hair$speciesID[hair$speciesID =='Mazama']<-'Ma'
hair$speciesID[hair$speciesID =='O. virginianus']<-'OV'
hair$speciesID[hair$speciesID =='P.bengalensis']<-'PB'
hair$speciesID[hair$speciesID =='P.hermaphroditus']<-'PH'
hair$speciesID[hair$speciesID =='P.lotor']<-'PL'
hair$speciesID[hair$speciesID =='P.vitulina']<-'PV'
hair$speciesID[hair$speciesID =='R.tarandus']<-'RT'
hair$speciesID[hair$speciesID =='R.Unicolor']<-'RU'
hair$speciesID[hair$speciesID =='T.ornatus']<-'TOP'
hair$speciesID[hair$speciesID =='T.ornatus.V']<-'TOV'
hair$speciesID[hair$speciesID =='U.americanus']<-'UA'
hair$speciesID[hair$speciesID =='U.arctos']<-'UR'
hair$speciesID[hair$speciesID =='H.malayanus']<-'HM'
hair$speciesID[hair$speciesID =='U.maritimus']<-'UM'
hair$speciesID[hair$speciesID =='U.thibetanus']<-'UT'
hair$speciesID[hair$speciesID =='V.vulpes']<-'VV'

hair<- droplevels(hair)
table(hair$speciesID)

#organize levels
hair$speciesID <- ordered(hair$speciesID,levels=c("CE","RT","OV","EC", "Ma","RU", "VV","PH","PL","EB","MF","MM", "CL","PB","LR","CT","PV", "UA","UT","UR","UM","HM","AM","MU","TOP","TOV"))
                          
hair$Diet <- ordered(hair$Diet,levels=c("herbivore","omnivore","carnivore","piscivore", "bear"))

```

#FUGURE 1

##Map
```{r}
library(ggplot2)
theme_set(theme_void() ) # setting background

world_map <- map_data("world")
a<-c(1:6)
b<-c("China","USA","Peru","Venezuela", "Indonesia","India")

countries_sampled<-data.frame(a,b)
names(countries_sampled) <- c("sample","country")

library(dplyr)
# join to dataset
map.world_joined <- left_join(world_map, countries_sampled, by = c('region' = 'country'))

map.world_joined <- map.world_joined %>% mutate(fill_flg = ifelse(is.na(sample),F,T))
head(map.world_joined)

map<-ggplot(map.world_joined, aes( x = long, y = lat, group = group )) +
  geom_polygon(aes(fill = fill_flg),colour = "white",alpha=0.5) +  scale_fill_manual(values = c("wheat","orange")) +    theme(legend.position = "none",  plot.background=element_rect(fill = "white")) +  coord_sf(ylim = c(-52, 80), xlim = c(-150, 170),datum = NA)
        
```

##subset data
```{r}
#subset data  for plot
Indonesia<- hair[ which(hair$region=='Borneo'),] 
Alaska<- hair[ which(hair$region=='Alaska'),] 
US<- hair[ which(hair$region=='New Mexico'),] 
China<- hair[ which(hair$region=='Szechuan'),]
India<- hair[ which(hair$region=='Jammu and Kashmir'),] 
SAmerica<- hair[ which(hair$region=='Peru and Venezuela'),] 
           
upper<-rbind(Alaska, US,China)

down<-rbind(SAmerica, India, Indonesia)
down$region <- ordered(down$region,levels=c("Peru and Venezuela","Jammu and Kashmir","Borneo")) #reordering levels

upper<- droplevels(upper)
down<- droplevels(down)

#set colors manually 
cols <- c("carnivore" = "#AF601A", "omnivore" = "#E59866", "herbivore" = "#A9DFBF", "bear" = "#ABB2B9","piscivore"="#BF360C")
```

##ensambling figure
```{r}
library(egg)
###upper figure
U_diet<-ggplot(upper, aes(speciesID, X15N,colour=Diet)) +  facet_wrap( ~ region, scales="free") +
stat_summary(fun.data = mean_se, geom = "pointrange",stat = "identity")  + 
  scale_color_manual(values = cols,labels=c("herbivore","omnivore","carnivore","piscivore","bear")) +
theme_classic() +
      labs(x="species", y = expression(paste(delta^15,"N"))) 

#There is only one point for M.ursinus so I'm adding it apart
F_upper<-U_diet + theme(legend.position="top",strip.background = element_rect(fill = "white", color = "white", size = 1),
  strip.text = element_text(colour = "black",size=11)) 

#get legend function # run first with legend.position="top" before to change to "none"
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

legend<- get_legend(F_upper)

#run again with legend
F_upper<-U_diet + theme(legend.position="none",strip.background = element_rect(fill = "white", color = "white", size = 1),
  strip.text = element_text(colour = "black",size=11)) 

####down figure 
D_diet<-ggplot(down, aes(speciesID, X15N,colour=Diet)) +  facet_wrap( ~ region, scales="free") +
stat_summary(fun.data = mean_se, geom = "pointrange",stat = "identity")  + 
  scale_color_manual(values = cols,labels=c("herbivore","omnivore","carnivore","piscivore","bear")) +
theme_classic()+  labs(x = "species", y = expression(paste(delta^15,"N")))

#There is only one point for M.ursinus so I'm adding it apart
F_down<- D_diet + theme(legend.position="none",strip.background = element_rect(fill = "white", color = "white", size = 1),
  strip.text = element_text(colour = "black",size=11)) + geom_point(data = hair[ which(hair$species=="M. ursinus"),], size=2.5,col = "#ABB2B9")

## Combine figures in one

grid.arrange(legend,F_upper, map, F_down, nrow = 4, ncol = 1,heights = c(0.2,1,1.5,1) )

# transparenci is not supported by EPS
 
cairo_ps(file = "map_N15.eps", onefile = FALSE, fallback_resolution = 600)
grid.arrange(legend,F_upper, map, F_down, nrow = 4, ncol = 1,heights = c(0.2,1,1.5,1) )
dev.off()
```

#JAGS
**Model: ~ SPECIES2 + REGION 

```{r}
hair2 <- read.csv(file="/Users/nereyda/Documents/Doctoral Thesis/Isotopes/hair_v2.csv", header=TRUE, sep=",")
M <- glm(X15N ~  species2+region, data = hair2) 
summary(M)
```

```{r}
######################################################

#Specifying the data for JAGs
X <- model.matrix(~ species2 +region,
                    data = hair2)
K <- ncol(X)

win.data <- list(Y    = hair2$X15N,
                 N    = nrow(hair2),
                 X    = X,
                 K   =  K
                 )

```

```{r}
#Jags modelling code
sink("Gauss.txt")
cat("
model{
  #2. Likelihood 
    for (i in 1:N){  
      Y[i] ~ dnorm(mu[i],tau)
      mu[i] <-inprod(beta[],X[i,])  
    }
      
    #1. Prior
 for (i in 1:K) {
   beta[i] ~ dnorm(0, 0.01)
 }
  sigma ~ dunif(0,10)
  tau <- 1/(sigma*sigma)
}

",fill = TRUE)

sink()
```

```{r}
#Set the initial values for the betas
inits <-function(){
  list(beta=rnorm(0.01,0,0.001),
       sigma=runif(1,0.001,1)
  )}
 
#Parameters to estimate

params <- c("beta","sigma")
```

```{r}

J <- jags(data = win.data,
           inits = inits,
           parameters = params,
           model.file = "Gauss.txt",
           n.thin = 10,
           n.chains = 3,
           n.burnin = 40000,
           n.iter   = 50000)

#save(J, file="M.RData")

##################

library(MCMCvis)

#load("M.RData")

out <- J$BUGSoutput
out<-print(out, intervals=c(0.025, 0.975), digits=2)

vars<-c("beta[1]","beta[2]","beta[3]", "beta[4]","beta[5]","beta[6]","beta[7]","beta[8]", "beta[9]","beta[10]","beta[11]","beta[12]","beta[13]", "beta[14]","beta[15]","beta[16]","beta[17]","beta[18]", "beta[19]","beta[20]","beta[21]","beta[22]","beta[23]", "beta[24]","beta[25]","beta[26]","beta[27]","beta[28]", "beta[29]","beta[30]","beta[31]")
OUT<- MyBUGSOutput(out,vars)

cbind(coef(M),OUT)

MCMCplot(J, 
         params = c("beta\\[1\\]","beta\\[2\\]","beta\\[3\\]", "beta\\[4\\]","beta\\[5\\]","beta\\[6\\]","beta\\[7\\]","beta\\[8\\]", "beta\\[9\\]","beta\\[10\\]","beta\\[11\\]","beta\\[12\\]","beta\\[13\\]", "beta\\[14]","beta\\[15\\]","beta\\[16\\]","beta\\[17\\]","beta\\[18\\]", "beta\\[19]","beta\\[20\\]","beta\\[21\\]","beta\\[22\\]","beta\\[23\\]", "beta\\[24\\]","beta\\[25\\]","beta\\[26\\]") , 
         
       labels = c('A.melanoleuca','C.elaphus', 'C.lupus', 'C.temminckii', 'E.Barbara' ,'E.cephalopus','L.rufus', 'M. ursinus',       'M.foina', 'M.moschata', 'Mazama', 'O. virginianus', 
                 'P.bengalensis', 'P.hermaphroditus', 'P.lotor', 'P.vitulina', 'R.tarandus' , 'R.Unicolor',
                 'T.ornatus.P', 'T.ornatus.V',  'U.americanus','U.arctos', 'H.malayanus','U.maritimus','U.thibetanus' ,  'V.vulpes'),
       rank = TRUE,
         horiz = FALSE,
       ISB =FALSE,
       ref_ovl = TRUE) 

MCMCplot(J, params = c("beta\\[1\\]","beta\\[29\\]","beta\\[31\\]","beta\\[30\\]", "beta\\[28\\]","beta\\[27\\]"),
          labels = c('Alaska-US','New Mexico-US','szechuan-China','Peru & Venezuela','Jammu and Kahmir-India','Borneo-Indonesia'),
         rank = TRUE,
         horiz = FALSE,
       ISB =FALSE,
       ref_ovl = TRUE) 

```

#posterior of bears   
```{r}

#Jags modelling code
sink("M_P1.txt")
cat("
model{
  #2. Likelihood 
    for (i in 1:N){  
      Y[i] ~ dnorm(mu[i],tau)
      mu[i] <-inprod(beta[],X[i,])  
    }
      
    #1. Prior
 for (i in 1:K) {
   beta[i] ~ dnorm(0, 0.01)
 }
  sigma ~ dunif(0,100)
  tau <- 1/(sigma*sigma)
  
  #predicted probabilities
  AM<- beta[1] + beta[31] 
  MU<-beta[1] + beta[8]+ beta[28] 
  URI<-beta[1] + beta[22] +beta[28] 
  HM<-beta[1] + beta[23] +beta[27] 
  UM<-beta[1] + beta[24] 
  UTI<-beta[1] + beta[25] +beta[28] 
  TOP<-beta[1] + beta[19] +beta[30] 
  TOV<-beta[1] + beta[20] +beta[30] 
  UAA<-beta[1] + beta[21] 
  UANM<-beta[1] + beta[21]+ beta[29]
  URA<-beta[1] + beta[22] 
  URNM<-beta[1] + beta[22]+ beta[29]
  UTC<-beta[1] + beta[25] +beta[31] 
pAM ~ dnorm(AM,tau)
pMU ~ dnorm(MU,tau)
pURI ~ dnorm(URI,tau)
pHM ~ dnorm(HM,tau)
pUM ~ dnorm(UM,tau)
pUTI ~ dnorm(UTI,tau)
pTOP ~ dnorm(TOP,tau)
pTOV ~ dnorm(TOV,tau)
pUAA ~ dnorm(UAA,tau)
pUANM ~ dnorm(UANM,tau)
pURA ~ dnorm(URA,tau)
pURNM ~ dnorm(URNM,tau)
pUTC ~ dnorm(UTC,tau)

}

",fill = TRUE)
sink()
```

```{r}
params <-c("pAM","pMU","pURI", "pHM", "pUM","pUTI", "pTOP", "pTOV", "pUAA" ,"pUANM","pURA", "pURNM", "pUTC")


JP <- jags(data = win.data,
           inits = inits,
           parameters = params,
           model.file = "M_P1.txt",
           n.thin = 10,
           n.chains = 3,
           n.burnin = 40000,
           n.iter   = 50000)

#save(JP, file="M_P1.RData")
#################

outP <- JP$BUGSoutput
outP<-print(outP, intervals=c(0.025, 0.975), digits=3)

vars<-c("pAM","pHM","pMU","pTOP", "pTOV", "pUAA" ,"pUANM", "pUM" ,"pURA", "pURI", "pURNM", "pUTC","pUTI")

MyBUGSHist(J$BUGSoutput,vars)
MyBUGSChains(JP$BUGSoutput,vars)
MyBUGSACF(JP$BUGSoutput,vars)

OUT_P1<- MyBUGSOutput(JP$BUGSoutput,vars)

```


#FIGURE 2
```{r}
oso<-as.data.frame(OUT_P1)
names(oso)[1] <- "mean"
name<-factor(c("A.melanoleuca","H.malayanus","M.ursinus","T.ornatus.P","T.ornatus.V",
              "U.americanus.A","U.americanus.NM","U.maritimus", "U.arctos.A", "U.arctos.I",
                "U.arctos.NM","U.thibetanus.C","U.thibetanus.I"))
oso$min<-oso$`2.5%`
oso$max<-oso$`97.5%`
#oso<-cbind(oso,sp)
#p <- ggplot(oso, aes(sp, mean, colour = sp))

sp2<-factor(c("AM","HM","MU", "TO.P","TO.V","UA.A","UA.NM","UM","UR.A","UR.I",
              "UR.NM","UT.C", "UT.I"
                   ))

oso<-cbind(oso,sp2,name)


##reordeing
p<-oso %>%
  arrange(mean) %>%
  mutate(sp2 = factor(sp2, levels = c("AM","TO.V","TO.P","UT.I" ,"UT.C", "UA.NM","UA.A",
  "UR.I","UR.NM","UR.A","HM","MU","UM"))) %>%
  ggplot(aes(x=sp2, y=mean,shape = name ))+ theme_classic()


#change color and symbol
col_oso<-c("AM"="#FAD7A0",
             "UT.I"="#F0B27A","UT.C"="#F0B27A",
             "TO.V"="#EB984E","TO.P"="#EB984E", 
             "UA.A"="#E59866","UA.NM"="#E59866", "HM"="#E59866",
  "UR.I"="#CA6F1E","UR.A"="#CA6F1E","UR.NM"="#CA6F1E","MU"="#BA4A00","UM"="#A04000")


p + geom_pointrange(aes(ymin = min, ymax = max,colour=sp2,fatten = 2), size=1.2) +
   scale_color_manual(values = col_oso,labels=c("AM","UT.I" ,"UT.C","TO.V","TO.P", "UA.A","UA.NM",
  "UL","UR.I","UR.A","UR.NM","MU","UM")) +
  theme(legend.position="none",axis.text=element_text(size=16),axis.title=element_text(size=17,face="bold")) + xlab("Bear species") + ylab( expression(paste("predicted",delta^15,"N"))) +
    scale_shape_manual(values = c(15,3,16,1,1,17,17,0,0,0,8,6,6)) +ylim(-8, 30)
```


##Pair comparison 
```{r}
#Jags modelling code
sink("M_DifBear.txt")
cat("
model{
  #2. Likelihood 
    for (i in 1:N){  
      Y[i] ~ dnorm(mu[i],tau)
      mu[i] <-inprod(beta[],X[i,])  
    }
      
    #1. Prior
 for (i in 1:K) {
   beta[i] ~ dnorm(0, 0.01)
 }
  sigma ~ dunif(0,100)
  tau <- 1/(sigma*sigma)
  
  #predicted probabilities
  B1<- beta[1] + beta[31] 
  B2<-beta[1] + beta[25] +beta[28] 
  B3<-beta[1] + beta[25] +beta[31] 
  B4<-beta[1] + beta[20] +beta[30] 
  B5<-beta[1] + beta[19] +beta[30] 
  B6<-beta[1] + beta[21] 
  B7<-beta[1] + beta[21]+ beta[29]
  B8<-beta[1] + beta[23] +beta[27] 
  B9<-beta[1] + beta[22] +beta[28] 
  B10<-beta[1] + beta[22]
  B11<-beta[1] + beta[22]+ beta[29]
  B12<-beta[1] + beta[8]+ beta[28]
  B13<-beta[1] + beta[24] 
  
  
#Diferrence 
D12<-B1-B2
D13<-B1-B3
D14<-B1-B4
D15<-B1-B5
D16<-B1-B6
D17<-B1-B7
D18<-B1-B8
D19<-B1-B9
D110<-B1-B10
D111<-B1-B11
D112<-B1-B12
D113<-B1-B13

D23<-B2-B3
D24<-B2-B4
D25<-B2-B5
D26<-B2-B6
D27<-B2-B7
D28<-B2-B8
D29<-B2-B9
D210<-B2-B10
D211<-B2-B11
D212<-B2-B12
D213<-B2-B13

D34<-B3-B4
D35<-B3-B5
D36<-B3-B6
D37<-B3-B7
D38<-B3-B8
D39<-B3-B9
D310<-B3-B10
D311<-B3-B11
D312<-B3-B12
D313<-B3-B13

D45<-B4-B5
D46<-B4-B6
D47<-B4-B7
D48<-B4-B8
D49<-B4-B9
D410<-B4-B10
D411<-B4-B11
D412<-B4-B12
D413<-B4-B13

D56<-B5-B6
D57<-B5-B7
D58<-B5-B8
D59<-B5-B9
D510<-B5-B10
D511<-B5-B11
D512<-B5-B12
D513<-B5-B13

D67<-B6-B7
D68<-B6-B8
D69<-B6-B9
D610<-B6-B10
D611<-B6-B11
D612<-B6-B12
D613<-B6-B13

D78<-B7-B8
D79<-B7-B9
D710<-B7-B10
D711<-B7-B11
D712<-B7-B12
D713<-B7-B13

D89<-B8-B9
D810<-B8-B10
D811<-B8-B11
D812<-B8-B12
D813<-B8-B13

D910<-B9-B10
D911<-B9-B11
D912<-B9-B12
D913<-B9-B13

D1011<-B10-B11
D1012<-B10-B12
D1013<-B10-B13

D1112<-B11-B12
D1113<-B11-B13

D1213<-B12-B13


  
pD12 ~ dnorm(D12,tau)
pD13 ~ dnorm(D13,tau)
pD14 ~ dnorm(D14,tau)
pD15 ~ dnorm(D15,tau)
pD16 ~ dnorm(D16,tau)
pD17 ~ dnorm(D17,tau)
pD18 ~ dnorm(D18,tau)
pD19 ~ dnorm(D19,tau)
pD110 ~ dnorm(D110,tau)
pD111 ~ dnorm(D111,tau)
pD112 ~ dnorm(D112,tau)
pD113 ~ dnorm(D113,tau)

pD23 ~ dnorm(D23,tau)
pD24 ~ dnorm(D24,tau)
pD25 ~ dnorm(D25,tau)
pD26 ~ dnorm(D26,tau)
pD27 ~ dnorm(D27,tau)
pD28 ~ dnorm(D28,tau)
pD29 ~ dnorm(D29,tau)
pD210 ~ dnorm(D210,tau)
pD211 ~ dnorm(D211,tau)
pD212 ~ dnorm(D212,tau)
pD213 ~ dnorm(D213,tau)

pD34 ~ dnorm(D34,tau)
pD35 ~ dnorm(D35,tau)
pD36 ~ dnorm(D36,tau)
pD37 ~ dnorm(D37,tau)
pD38 ~ dnorm(D38,tau)
pD39 ~ dnorm(D39,tau)
pD310 ~ dnorm(D310,tau)
pD311 ~ dnorm(D311,tau)
pD312 ~ dnorm(D312,tau)
pD313 ~ dnorm(D313,tau)

pD45 ~ dnorm(D45,tau)
pD46 ~ dnorm(D46,tau)
pD47 ~ dnorm(D47,tau)
pD48 ~ dnorm(D48,tau)
pD49 ~ dnorm(D49,tau)
pD410 ~ dnorm(D410,tau)
pD411 ~ dnorm(D411,tau)
pD412 ~ dnorm(D412,tau)
pD413 ~ dnorm(D413,tau)

pD56 ~ dnorm(D56,tau)
pD57 ~ dnorm(D57,tau)
pD58 ~ dnorm(D58,tau)
pD59 ~ dnorm(D59,tau)
pD510 ~ dnorm(D510,tau)
pD511 ~ dnorm(D511,tau)
pD512 ~ dnorm(D512,tau)
pD513 ~ dnorm(D513,tau)

pD67 ~ dnorm(D67,tau)
pD68 ~ dnorm(D68,tau)
pD69 ~ dnorm(D69,tau)
pD610 ~ dnorm(D610,tau)
pD611 ~ dnorm(D611,tau)
pD612 ~ dnorm(D612,tau)
pD613 ~ dnorm(D613,tau)

pD78 ~ dnorm(D78,tau)
pD79 ~ dnorm(D79,tau)
pD710 ~ dnorm(D710,tau)
pD711 ~ dnorm(D711,tau)
pD712 ~ dnorm(D712,tau)
pD713 ~ dnorm(D713,tau)

pD89 ~ dnorm(D89,tau)
pD810 ~ dnorm(D810,tau)
pD811 ~ dnorm(D811,tau)
pD812 ~ dnorm(D812,tau)
pD813 ~ dnorm(D813,tau)

pD910 ~ dnorm(D910,tau)
pD911 ~ dnorm(D911,tau)
pD912 ~ dnorm(D912,tau)
pD913 ~ dnorm(D913,tau)

pD1011 ~ dnorm(D1011,tau)
pD1012 ~ dnorm(D1012,tau)
pD1013 ~ dnorm(D1013,tau)

pD1112 ~ dnorm(D1112,tau)
pD1113 ~ dnorm(D1113,tau)

pD1213 ~ dnorm(D1213,tau)

}

",fill = TRUE)
sink()

```

```{r}

params <-c("pD12","pD13","pD14", "pD15", "pD16","pD17", "pD18","pD19", "pD110" ,"pD111","pD112","pD113",
           "pD23","pD24", "pD25", "pD26","pD27", "pD28","pD29", "pD210" ,"pD211","pD212", "pD213",
           "pD34", "pD35", "pD36","pD37", "pD38","pD39", "pD310" ,"pD311","pD312", "pD313",
           "pD45", "pD46","pD47", "pD48","pD49", "pD410" ,"pD411","pD412", "pD413",
           "pD56","pD57", "pD58","pD59", "pD510" ,"pD511","pD512", "pD513",
           "pD67", "pD68","pD69", "pD610" ,"pD611","pD612", "pD613",
           "pD78","pD79", "pD710" ,"pD711","pD712", "pD713",
           "pD89", "pD810" ,"pD811","pD812", "pD813",
           "pD910" ,"pD911","pD912", "pD913",
           "pD1011","pD1012", "pD1013",
           "pD1112", "pD1113",
           "pD1213")


J_Dif <- jags(data = win.data,
           inits = inits,
           parameters = params,
           model.file = "M_DifBear.txt",
           n.thin = 10,
           n.chains = 3,
           n.burnin = 40000,
           n.iter   = 50000)

#save(J_Dif, file="M_Dif.RData")

#################


outD <- J_Dif$BUGSoutput
outD<-print(outD, intervals=c(0.025, 0.975), digits=2)


 MCMCplot(J_Dif, 
         params = params,
           labels = c("1-2","pD13","pD14", "pD15", "pD16","pD17", "pD18","pD19", "pD110" ,"pD111","pD112","pD113",
           "pD23","pD24", "pD25", "pD26","pD27", "pD28","pD29", "pD210" ,"pD211","pD212", "pD213",
           "pD34", "pD35", "pD36","pD37", "pD38","pD39", "pD310" ,"pD311","pD312", "pD313",
           "pD45", "pD46","pD47", "pD48","pD49", "pD410" ,"pD411","pD412", "pD413",
           "pD56","pD57", "pD58","pD59", "pD510" ,"pD511","pD512", "pD513",
           "pD67", "pD68","pD69", "pD610" ,"pD611","pD612", "pD613",
           "pD78","pD79", "pD710" ,"pD711","pD712", "pD713",
           "pD89", "pD810" ,"pD811","pD812", "pD813",
           "pD910" ,"pD911","pD912","pD913",
           "pD1011","pD1012","pD1013", 
           "pD1112", "pD1113",
           "pD1213"),
        sz_labels=0.7,
        sz_med=1,
        sz_thick=2,
        sz_thin=1,
        xlab="Posterior differences",
  sz_ax_txt=1.2,      
sz_tick_txt=1,
       ref_ovl = TRUE) 
        
```

#posterior all
```{r}
##M19: species2+region

#Jags modelling code
sink("M_PT.txt")
cat("
model{
  #2. Likelihood 
    for (i in 1:N){  
      Y[i] ~ dnorm(mu[i],tau)
      mu[i] <-inprod(beta[],X[i,])  
    }
      
    #1. Prior
 for (i in 1:K) {
   beta[i] ~ dnorm(0, 0.01)
 }
  sigma ~ dunif(0,10)
  tau <- 1/(sigma*sigma)
  
  #predicted probabilities
  AM<- beta[1] + beta[31] 
  MU<-beta[1] + beta[8]+ beta[28] 
  URI<-beta[1] + beta[22] +beta[28] 
  HM<-beta[1] + beta[23] +beta[27] 
  UM<-beta[1] + beta[24] 
  UTI<-beta[1] + beta[25] +beta[28] 
  TOP<-beta[1] + beta[19] +beta[30] 
  TOV<-beta[1] + beta[20] +beta[30] 
  UAA<-beta[1] + beta[21] 
  UANM<-beta[1] + beta[21]+ beta[29]
  URA<-beta[1] + beta[22] 
  URNM<-beta[1] + beta[22]+ beta[29]
  UTC<-beta[1] + beta[25] +beta[31] 
  
CEL<-beta[1] + beta[2] +beta[28]
CLA<-beta[1] + beta[3] 
CTC<-beta[1] + beta[4] +beta[31]
EB <-beta[1] + beta[5] +beta[30]
ECC<-beta[1] + beta[6] +beta[31]
LRN<-beta[1] + beta[7] +beta[19]
MFO<-beta[1] + beta[9] +beta[28]
MMC<-beta[1] + beta[10] +beta[31]
MA <-beta[1] + beta[11] +beta[30]
OVN<-beta[1] + beta[12] +beta[19]
PBI<-beta[1] + beta[13] +beta[28]
PBB<-beta[1] + beta[13] +beta[27]
PHB<-beta[1] + beta[14] +beta[27]
PLN<-beta[1] + beta[15] +beta[19]
PVI<-beta[1] + beta[16] 
RTA<-beta[1] + beta[17] 
RUB<-beta[1] + beta[18] +beta[27]
VVA<-beta[1] + beta[26] 

  
pAM ~ dnorm(AM,tau)
pMU ~ dnorm(MU,tau)
pURI ~ dnorm(URI,tau)
pHM ~ dnorm(HM,tau)
pUM ~ dnorm(UM,tau)
pUTI ~ dnorm(UTI,tau)
pTOP ~ dnorm(TOP,tau)
pTOV ~ dnorm(TOV,tau)
pUAA ~ dnorm(UAA,tau)
pUANM ~ dnorm(UANM,tau)
pURA ~ dnorm(URA,tau)
pURNM ~ dnorm(URNM,tau)
pUTC ~ dnorm(UTC,tau)

pCEL ~ dnorm(CEL,tau)
pCLA ~ dnorm(CLA,tau)
pCTC ~ dnorm(CTC,tau)
pEB ~ dnorm(EB,tau)
pECC ~ dnorm(ECC,tau)
pLRN ~ dnorm(LRN,tau)
pMFO ~ dnorm(MFO,tau)
pMMC ~ dnorm(MMC,tau)
pMA ~ dnorm(MA,tau)
pOVN ~ dnorm(OVN,tau)
pPBI ~ dnorm(PBI,tau)
pPBB ~ dnorm(PBI,tau)
pPHB ~ dnorm(PHB,tau)
pPLN ~ dnorm(PLN,tau)
pPVI ~ dnorm(PVI,tau)
pRTA ~ dnorm(RTA,tau)
pRUB ~ dnorm(RUB,tau)
pVVA ~ dnorm(VVA,tau)
}

",fill = TRUE)
sink()

```

```{r}

params <-c("pAM","pMU","pURI", "pHM", "pUM","pUTI", "pTOP", "pTOV", "pUAA" ,"pUANM","pURA", "pURNM", "pUTC","pCEL","pCLA","pCTC","pEB","pECC","pLRN","pMFO","pMMC","pMA","pOVN","pPBI","pPBB","pPHB","pPLN","pPVI",
           "pRTA","pRUB","pVVA")

J_PT <- jags(data = win.data,
           inits = inits,
           parameters = params,
           model.file = "M_PT.txt",
           n.thin = 10,
           n.chains = 3,
           n.burnin = 40000,
           n.iter   = 50000)

#save(J_PT, file="M_PT.RData")

#################

#TABLE1

outPT <- J_PT$BUGSoutput
outPT<-print(outPT, intervals=c(0.025, 0.975), digits=2)
```

## Literature data

ombine Means and SDs of bear population into one Group (bear species)
```{r}
# function from
# https://training.cochrane.org/handbook/current/chapter-06#section-6-5-2


# Section 2: First option: StatsToDo formula

Calc1 <- function(myDat) # myDat is matrix with 3 cols of n, mean, and SD
{
  m = nrow(myDat)  # number of groups
  tn = 0
  tx = 0
  txx = 0
  for(i in 1:m)
  {
    n = myDat[i,1]
    mean = myDat[i,2]
    sd = myDat[i,3]
    x = n * mean
    xx = sd^2*(n - 1) + x^2 / n 
    out<-cat("grp",i," n=",n," mean=",mean," SD=", sd, " Ex=", x, " Exx=",xx, "\n")
    tn = tn + n
    tx = tx + x
    txx = txx + xx
  }
  tmean = tx / tn
  tsd = sqrt((txx - tx^2/tn) / (tn - 1))
  out <- cat("Combined","n=",tn," mean=",tmean," SD=", tsd, " Ex=", tx, " Exx=",txx,"\n")
  c(tn,tmean,tsd)
}
```

###creating new dataset 
```{r}
library(tidyverse)
oso_N<-oso %>% select(1, 2,8)

#write.csv(oso_N, file = "oso_N.csv")
#/Users/nereyda/Dropbox/isotopos/oso_N.csv

#fusion U. thibetanus
g1 <- c(5, 4.340989, 1.673816) #n mean,sd of grp 1
g2 <- c(5, 3.420649, 1.695043) #n mean,sd of grp 2

Ut <- Calc1(matrix(data=c(g1,g2), ncol=3, byrow=TRUE)) # combine 3 groups entry data into matrix and call function
#10.000000  3.880819  1.660552

#fusion T. ornatus
g1 <- c(6, 4.498520, 1.643286) #n mean,sd of grp 1
g2 <- c(4, 4.313348, 1.724513) #n mean,sd of grp 2

To <- Calc1(matrix(data=c(g1,g2), ncol=3, byrow=TRUE)) 
#10.000000  4.424451  1.581352

#fusion U. americanus
g1 <- c(3, 4.243162, 1.664283) #n mean,sd of grp 1
g2 <- c(6, 5.982400, 1.643405) #n mean,sd of grp 2

Ua <- Calc1(matrix(data=c(g1,g2), ncol=3, byrow=TRUE)) 
#9.000000 5.402654 1.771069

#fusion U. arctos
g1 <- c(8, 6.666231, 1.616619) #n mean,sd of grp 1
g2 <- c(2, 1.522400, 1.828247) #n mean,sd of grp 2
g3 <- c(8, 8.324409, 1.574946) #n mean,sd of grp 2

Ur <- Calc1(matrix(data=c(g1,g2,g3), ncol=3, byrow=TRUE)) 
#18.000000  6.831662  2.582995

#create new dataframe
name<-factor(c("A.melanoleuca","U.thibetanus","T.ornatus", "U.americanus","H.malayanus","U.arctos","M.ursinus","U.maritimus"))
mean<-c(1.535481, 3.880819, 4.424451, 5.402654, 6.270083, 6.831662,6.753491, 21.056380)
sd_N <- c(1.628057, 1.660552, 1.581352,1.771069,1.708794, 2.582995, 2.146748 , 1.676094)

oso_N15<-data.frame(name, mean,sd_N)

```

#work with Literature review data and create dummy variables
```{r}
#diet <- read.csv(file="/Users/nereyda/Dropbox/Doctoral Thesis/Isotopes/R_diet.csv", header=TRUE, sep=",") #raw
diet2 <- read.csv(file="/Users/nereyda/Dropbox/Doctoral Thesis/Isotopes/diet2.csv", header=TRUE, sep=",") 

diet2$name<-diet2$bear
#subset
veg <- diet2[diet2$Food == "vegetal", ]
ins<- diet2[diet2$Food == "insect", ]
vrt<- diet2[diet2$Food == "vertebrate", ]

veg<-veg$len
ins<-ins$len
vrt<-vrt$len 
Lit<-cbind(oso_N15,veg,ins,vrt)

library(corrplot)

dat<-Lit[,2:6]
corrplot(cor(dat),
  method = "number",
  type = "upper" # show only upper side
)

# Pearson correlation test
veg <- cor.test(Lit$mean, Lit$veg)
#-0.7883103  #t = -3.1384, df = 6, p-value = 0.02011
vrt <- cor.test(Lit$mean, Lit$vrt)
#0.9333812  #t = 6.3706, df = 6, p-value = 0.0007027
ins <- cor.test(Lit$mean, Lit$ins)
#-0.1577879  #t = -0.3914, df = 6, p-value = 0.709

```

# include conflict lit rev
```{r}
sum <- read.csv(file="/Users/nereyda/Documents/Doctoral Thesis/Isotopes/summary.csv")
pred<-sum$I_pred
crop<-sum$I_crop

conf<-cbind(Lit,pred, crop)
# Pearson correlation test predation......no
cor.test(conf$pred, conf$mean)
#-0.3169812 #t = -0.81866, df = 6, p-value = 0.4443
cor.test(conf$pred, conf$veg)
#0.370535  #t = 0.97718, df = 6, p-value = 0.3662
cor.test(conf$pred, conf$vrt)
#-0.26936  #t = -0.68512, df = 6, p-value = 0.5188

# Pearson correlation test crops......no
cor.test(conf$crop, conf$mean)
#-0.1876662  #t = -0.468, df = 6, p-value = 0.6563
cor.test(conf$crop, conf$veg)
#-0.2379825   #t = -0.60018, df = 6, p-value = 0.5703
cor.test(conf$crop, conf$vrt)
#-0.4020061  #t = -1.0754, df = 6, p-value = 0.3235

```
